final-project

### 2024-11-22 code/final-project.Rmd
table 3 updates

```{r}
# Figure 2: Log-log plots to check Proportional Hazards assumption
library(gridExtra)
library(survminer)
library(survival)

plot_subtype <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ subtype), 
  data = seer, 
  fun ="cloglog",
  ylab = "Log-log survival probability",
  xlab = "Time (months)",
  legend.labs = c("Luminal A", "Luminal B", "HER2 Positive", "Triple Negative"))

plot_agecat <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ agecat), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("20-49 years", "50-69 years", ">= 70 years"))

plot_race <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ racecat), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("White", "Black", "Other"))

plot_marriage <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ marriage_status), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("Unmarried", "Married"))

plot_n_stage <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ n_stage), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("N0/1", "N2/3"))

plot_m_stage <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ m_stage), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("M0", "M1"))

plot_grade <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ grade), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("I/II", "III/IV"))

plot_radiation <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ radiation), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("No/Unknown", "Yes"))

plot_chemotherapy <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ chemotherapy), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("No/Unknown", "Yes"))

plot_surgery <- ggsurvplot(
  survfit(Surv(time = survival_months, event = survival_status) ~ surgery), 
  data = seer, 
  fun="cloglog",
  ylab= "Log-log survival probability",
  xlab= "Time (months)",
  legend.labs = c("No/Unknown", "Yes"))

grid.arrange(plot_subtype$plot, plot_agecat$plot, plot_race$plot, plot_marriage$plot, plot_n_stage$plot,
             plot_m_stage$plot, plot_grade$plot, plot_radiation$plot, plot_chemotherapy$plot, plot_surgery$plot,
             ncol = 5, nrow = 2)
```

```{r}
# check the proportional hazards assumption for each covariate with (1) a log log plot

## subtype
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ subtype), data = seer, fun="cloglog")

## agecat
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ agecat), data = seer, fun="cloglog")

## race
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ racecat), data = seer, fun="cloglog")

## marriage status
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ marriage_status), data = seer, fun="cloglog")

## N stage
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ n_stage), data = seer, fun="cloglog")

## M stage
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ m_stage), data = seer, fun="cloglog")

## grade
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ grade), data = seer, fun="cloglog")

## radiation
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ radiation), data = seer, fun="cloglog")

## chemotherapy
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ chemotherapy), data = seer, fun="cloglog")

## surgery
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ surgery), data = seer, fun="cloglog")
```

@@

```{r}
model <- coxph(
  Surv(survival_months, survival_status) ~ 
    relevel(factor(subtype), ref = '0') +
    relevel(factor(agecat), ref = '0') + 
    relevel(factor(racecat), ref = '0') +
    relevel(factor(marriage_status),ref='1') + 
    relevel(factor(n_stage),ref='0') + 
    relevel(factor(m_stage),ref='0') +
    relevel(factor(grade), ref='0') +
    relevel(factor(radiation), ref='1') + 
    relevel(factor(chemotherapy), ref='1') +
    relevel(factor(surgery), ref='1'),
  data = seer, 
  method = "breslow")

@@

@@

cor <- apply(resid$y, 2, \(x) {
  test <- cor.test(rank(resid$time), x)  # Perform correlation test
  c(
    correlation = test$estimate,
    p_value = test$p.value,
    conf_low = test$conf.int[1],
    conf_high = test$conf.int[2]
  ) 
})

cor_df <- as.data.frame(t(cor))
colnames(cor_df) <- c("Correlation", "P_Value", "Conf_Low", "Conf_High")
rownames(cor_df) <- c(
  "subtype1",
  "subtype2",
  "subtype3",
  "agecat1",
  "agecat2",
  "racecat1",
  "racecat2",
  "marraige_status0",
  "n_stage1",
  "m_stage",
  "grade",
  "radiation",
  "chemotherapy",
  "surgery"
)

>>>>

@@

```{r}
# check the correlation between schoenfeld residuals and ranked survival times

# run the cox ph model
model <- coxph(
  Surv(survival_months, survival_status) ~ 
    relevel(factor(subtype), ref = '0') +
    relevel(factor(agecat), ref = '0') + 
    relevel(factor(racecat), ref = '0') +
    relevel(factor(marriage_status),ref='1') + 
    relevel(factor(n_stage),ref='0') + 
    relevel(factor(m_stage),ref='0') +
    relevel(factor(grade), ref='0') +
    relevel(factor(radiation), ref='1') + 
    relevel(factor(chemotherapy), ref='1') +
    relevel(factor(surgery), ref='1'),
  data = seer, 
  method = "breslow")

@@

options(scipen=999)

@@

cor <- apply(resid$y, 2, \(x) {
  test <- cor.test(rank(resid$time), x)  # Perform correlation test
  c(
    correlation = round(test$estimate, 3),
    p_value = round(test$p.value, 3),
    conf_low = round(test$conf.int[1], 3),
    conf_high = round(test$conf.int[2], 3)
  ) 
})

# Convert the matrix to a data frame
cor_df_all <- as.data.frame(t(cor))

# Now, we can correctly mutate and format the output
cor_df <- cor_df_all %>%
  mutate(
    Correlation = paste(correlation.cor, " (", conf_low, ", ", conf_high, ")", sep = ""),
    P_Value = as.character(p_value)
  ) %>%
  select(Correlation, P_Value)
rownames(cor_df) <- c(
  "subtype1",
  "subtype2",
  "subtype3",
  "agecat1",
  "agecat2",
  "racecat1",
  "racecat2",
  "marraige_status0",
  "n_stage1",
  "m_stage",
  "grade",
  "radiation",
  "chemotherapy",
  "surgery"
)

# table 3: Correlation between Schoenfeld residuals and ranked survival time
t3_raw <- cor_df %>%
  add_row(Correlation = "", P_Value = "", .before = 1) %>%
  add_row(Correlation = "", P_Value = "", .before = 1) %>%
  add_row(Correlation = "", P_Value = "", .before = 6) %>%
  add_row(Correlation = "", P_Value = "", .before = 6) %>%
  add_row(Correlation = "", P_Value = "", .before = 10) %>%
  add_row(Correlation = "", P_Value = "", .before = 10) %>%
  add_row(Correlation = "", P_Value = "", .before = 14) %>%
  add_row(Correlation = "", P_Value = "", .before = 14) %>%
  add_row(Correlation = "", P_Value = "", .before = 17) %>%
  add_row(Correlation = "", P_Value = "", .before = 17) %>%
  add_row(Correlation = "", P_Value = "", .before = 20) %>%
  add_row(Correlation = "", P_Value = "", .before = 20) %>%
  add_row(Correlation = "", P_Value = "", .before = 23) %>%
  add_row(Correlation = "", P_Value = "", .before = 23) %>%
  add_row(Correlation = "", P_Value = "", .before = 26) %>%
  add_row(Correlation = "", P_Value = "", .before = 26) %>%
  add_row(Correlation = "", P_Value = "", .before = 29) %>%
  add_row(Correlation = "", P_Value = "", .before = 29) %>%
  add_row(Correlation = "", P_Value = "", .before = 32) %>%
  add_row(Correlation = "", P_Value = "", .before = 32)
